Automatic Identification and Annotation of Tasks in Structured Programs

### 2018-01-23 docs/ACM/paper.tex
Adding discussion about clauses.

\paragraph{The annotation zoo.}
Our equipment of choice to reveal task parallelism in programs is OpenMP 3.0.
This annotation system provides us with a rich set of pragmas that we can use
to indicate to compiler and runtime environment which tasks must be created.
We describe below which annotations we are using, and explain, informally, their
semantics.
A full overview of the syntax and semantics of these annotations is publicly
available\footnote{For a quick overview, we refer the reader to the leaflet
``Summary of OpenMP 3.0 C/C++ Syntax", which is made available by the OpenMP
group.}.

\begin{compactitem}
\item \textsf{task} {\em clauses}: creates a new task, according to the
parameters specified in the {\em clauses} list:
\begin{compactitem}
\item \textsf{default([shared/private])}: indicates that variables are either
shared or replicated among tasks.
\item \textsf{firstprivate($v$)}: indicates that $v$ must be replicated among
tasks, and initialized with its value at the point where the annotation is
executed.
\item \fer{Pedro/Gleison: can you guys continue this list?}
\end{compactitem}
\item \textsf{taskwait}: forces tasks to synchronize at the point of annotation.
\end{compactitem}

>>>>

\paragraph{The annotation zoo.}
Our equipment of choice to reveal task parallelism in programs is OpenMP 3.0.
This annotation system provides us with a set of pragmas that indicate to compiler
and runtime environment when and how tasks must be created.
We describe below which annotations we are using, and explain, informally, their
semantics.
Full overviews of the syntax and semantics of these annotations is publicly
available\footnote{For a quick overview, we refer the reader to the leaflet
``Summary of OpenMP 3.0 C/C++ Syntax", which is made available by the OpenMP
group.}; hence, we shall not dive into their details.
%
\begin{compactitem}
\item \textsf{parallel} (\textit{clauses}): forms a team of threads
which will execute the marked program region in parallel.
\item \textsf{single} (\textit{clauses}): specifies that a
program region must be executed by one thread in the team.
\item \textsf{task} (\textit{clauses}): creates a new task.
\item \textsf{taskwait}: defines a synchronization point where
threads must wait for the completion of child tasks.
\end{compactitem}
%
Some pragmas are associated with lists of clauses.
There are several of these clauses available in OpenMP 3.0; however, we only use
the following:
\begin{compactitem}
\item \textsf{default([shared/private])}: indicates that variables are either
shared or replicated among tasks.
\item \textsf{firstprivate($v$)}: indicates that $v$ must be replicated among
tasks, and initialized with its value at the point where the annotation is
executed.
\item \textsf{untied}: if a task has this modifier, then its associated code can
be executed by more than one thread; e.g., threads might alternate execution due
to preemption.
\item \textsf{depend}($\mathit{in}/\mathit{out}/\mathit{inout}$):
determine if data is read ($\mathit{in}$), written ($\mathit{out}$) or both
within a task region; thus, effectively setting dependences among tasks.
\item \textsf{depend}({\em condition}): this clause defines the conditions
under which a task is allowed to create threads.
We use it to enforce the cost model.
\end{compactitem}
